## Exercise 9: For loop

Create the script "exercise9.R" and save it to the "Rcourse/Module2" directory: you will save all the commands of exercise 9 in that script.
<br>Remember you can comment the code using #.


<details>
<summary>
correction
</summary>

```{r, eval=F}
getwd()
setwd("~/Rcourse/Module2")
```

</details>


**1- Write a for loop that iterates over 2 to 10 and prints the square root of each number (function sqrt()).**

<details>
<summary>
correction
</summary>

```{r}
for(i in 2:10){
	print(sqrt(i))
}
```

</details>


**2- Write a for loop that iterates over 5 to 15 and prints a vector of 2 elements containing each number and its square root**

<details>
<summary>
correction
</summary>

```{r}
for(i in 5:15){
	veci <- c(i, sqrt(i))
        print(veci)
}
```

</details>

**3- Create the following matrix**

```{r}
mat1 <- matrix(rnorm(40), nrow=20)
```

* Write a for loop that iterates over each row of mat1 and prints the median value of each row.

<details>
<summary>
correction
</summary>

```{r}
for(j in 1:nrow(mat1)){
	# extract the row
	rowj <- mat1[j,]
	# print rowj
	print(rowj)
	# print median value in row
	print(median(rowj))
}       
```     

</details>

