# Data filtering

We will now see a couple of basic functions to filter and select a data object.

## filter()

**filter()** will filter **rows**.

If we take the **geneexp** objet as an example, we can filter only down-regulated genes using the **filter()** function of **{dplyr}**.

In case you need to import the data again:

```{r, eval=FALSE}
geneexp <- read_csv("DataViz_source_files-main/files/expression_20genes.csv")
```

```{r}
filter(geneexp, DE=="Down")
```

**==** is a **logical operator** that represens equality. It means that filter will return **rows in geneexp that are exactly equal to "Down"**.

Logical operators:

| Operator  | Description   |
|:--------: |:-------------:|
|     <     | less than    |
|     <=    | less than or equal to   |
|     >     | greater than  |
|     >=    | greater than or equal to|
|     ==    | exactly equal to   |
|     !=    | not equal to |
|    !x     | not x |
|    x | y  | x OR y |
|    x & y  | x AND y |

We can have several conditions.

For example, we may want to extract only **Up** or **Down** rows from geneexp column **DE**:

```{r}
filter(geneexp, DE=="Down" | DE=="Up")
```

Here, we introduce another operator, **|**, which mean **OR**, so rows will be kept if there is either **Down** or **Up** in **DE** column.

A good practice is to assign the filtered output to a new object, for example:

```{r}
geneexp_filt <- filter(geneexp, DE=="Down" | DE=="Up")
```

## select()

**select()** will select **columns**.

Select only columns **Gene** and **DE**:

```{r}
select(geneexp, Gene, DE)
```


## Exercise




