# More exercises

The following sets of exercises aim to help you practice what was learned during this workshop.
<br>

Exercise 5: barplot
<br>
Exercise 6: scatter plot
<br>
Exercise 7: boxplot
<br>

## Exercise 5: barplots

1. Import **DataViz_source_files-main/files/population_age_group_per_sex_long_format.csv** into an object called **pop_age**.

<details>
<summary>
correction
</summary>

```{r}
pop_age <- read_csv("DataViz_source_files-main/files/population_age_group_per_sex_long_format.csv")
```

</details>

<br>

2. Create a barplot that shows the number of people per **Year**:

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=pop_age, mapping=aes(x=Year, y=Population)) + 
  geom_bar(stat="identity")
```

</details>

<br>

3. Split the view by Sex:


<details>
<summary>
correction
</summary>

```{r}
ggplot(data=pop_age, mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity")
```

</details>

<br>

4. Select only "Spain" in "Country":

  
<details>
<summary>
correction
</summary>

```{r}
ggplot(data=filter(pop_age, Country=="Spain"), mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity")
```

or:

```{r, eval=F}
filter(pop_age, Country=="Spain") %>% ggplot(mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity")
```

</details>

<br>

5. Show the bars next to each other.

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=filter(pop_age, Country=="Spain"), mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity", position="dodge")
```

</details>

<br>


6. Change default colors using **scale_fill_manual()** layer.

The structure is not easy to remember: take a peek at the answer!

  
<details>
<summary>
correction
</summary>

```{r}
ggplot(data=filter(pop_age, Country=="Spain"), mapping=aes(x=Year, y=Population, fill=Sex)) + 
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=c(Female="darkgreen", Male="red"))
```

</details>

<br>

7. Add a title, change the theme / background:

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=filter(pop_age, Country=="Spain"), mapping=aes(x=Year, y=Population, fill=Sex)) +
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=c("darkgreen", "red")) + 
  ggtitle("Spanish population from 1950 to 2021 per sex") + 
  theme_minimal()
```

</details>

<br>
8. Now split the plot by **Age group** using **facet_wrap()**

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=filter(pop_age, Country=="Spain"), mapping=aes(x=Year, y=Population, fill=Sex)) +
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=c("darkgreen", "red")) + 
  ggtitle("Spanish population from 1950 to 2021 per sex") + 
  theme_minimal() +
  facet_wrap(~Age_group)
```

</details>
<br>

9. Save to a PDF file: Export -> Save as PDF. Note: when saving to PDF format, you cannot manualy drag the plot to get the size and proportion you want, but you can **Preview** before saving it.

If you prefer, you can save the file using **ggsave()** (or you can do both!).

<details>
<summary>
correction
</summary>

```{r}
# first, save to an object
plot_age <- ggplot(data=filter(pop_age, Country=="Spain"), mapping=aes(x=Year, y=Population, fill=Sex)) +
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=c("darkgreen", "red")) + 
  ggtitle("Spanish population from 1950 to 2021 per sex") + 
  theme_minimal() +
  facet_wrap(~Age_group)

# save file
ggsave(filename="population_age_sex.pdf", plot=plot_age, device = "pdf", units = "in", width = 10, height = 8)
```

</details>

<br>

10. Save your filtered data: have a look at the **write_csv()** function:

<details>
<summary>
correction
</summary>

```{r}
# first, save to an object
mysubset <- filter(pop_age, Country=="Spain")
write_csv(mysubset, "subset_spain.csv")
```

</details>

## Exercise 6: scatter plot

1. Import **DataViz_source_files-main/files/GSE150029_rnaseq_log2.csv** into an object called **rnaseq**.

<details>
<summary>
correction
</summary>

```{r}
rnaseq <- read_csv("DataViz_source_files-main/files/GSE150029_rnaseq_log2.csv")
```

</details>
<br>

2. Create a scatter plot that represents sample CTRL on the x axis and sample EZH on the y axis.

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=rnaseq, mapping=aes(x=CTRL, y=EZH)) +
  geom_point()
```

</details>

<br>
3. Color the points according to the **gene_biotype**

<details>
<summary>
correction
</summary>

```{r, fig.width=16}
ggplot(data=rnaseq, mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point()
```

</details>

<br>
3. Not very readable! Filter and plot only data corresponding to either lincRNA OR miRNA.

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=filter(rnaseq, gene_biotype=="lincRNA" | gene_biotype=="miRNA"), mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point()

# using the pipe:
filter(rnaseq, gene_biotype=="lincRNA" | gene_biotype=="miRNA") %>% ggplot(mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point()
```

</details>

<br>
4. Now select and only those lincRNAs and miRNAs that are expressed in CTRL at least 1.5 times more than in EZH.

<details>
<summary>
correction
</summary>

```{r}
filter(rnaseq, gene_biotype=="lincRNA" | gene_biotype=="miRNA" & CTRL > 1.5*EZH) %>% ggplot(mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point()
```

</details>

5. Add a title to the plot, and make it **bold** (see **theme()** section of the course)

<details>
<summary>
correction
</summary>

```{r}
filter(rnaseq, gene_biotype=="lincRNA" | gene_biotype=="miRNA" & CTRL > 1.5*EZH) %>% ggplot(mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point() +
  ggtitle("lincRNA and miRNA") +
  theme(plot.title = element_text(face = "bold"))
```

</details>
<br>

NOTE: If you want to label only one (or few) point(s), you can do it the following way:

First, filter the data frame:
```{r}
SNHG8 <- filter(rnaseq, gene_name=="SNHG8")
```

Then, add it to geom_text:

```{r}
filter(rnaseq, gene_biotype=="lincRNA" | gene_biotype=="miRNA" & CTRL > 1.5*EZH) %>% ggplot(mapping=aes(x=CTRL, y=EZH, color=gene_biotype)) +
  geom_point() +
  ggtitle("lincRNA and miRNA") +
  theme(plot.title = element_text(face = "bold")) +
  geom_text(data=SNHG8, label="SNHG8", show.legend = FALSE)
```



## Exercise 7: boxplot

1. Convert **DataViz_source_files-main/files/GSE150029_rnaseq_log2.csv** (rnaseq object) to a long format, and save it as a new object called **rnaseq2**.

Note that it exists already as a separate object, if you prefer to read this in directly: **DataViz_source_files-main/files/GSE150029_rnaseq_log2_long.csv**.

<details>
<summary>
correction
</summary>

```{r}
rnaseq2 <- rnaseq %>% pivot_longer(cols=c(CTRL, EZH), names_to = "sample", values_to = "log2_counts")
```

</details>

<br>

2. Create a boxplot that will represent the samples on the x axis, and their expression on the y axis.

<details>
<summary>
correction
</summary>

```{r}
ggplot(data=rnaseq2, mapping=aes(x=sample, y=log2_counts)) + 
  geom_boxplot()
```

</details>

<br>

3. Split the boxes per gene_biotype.

<details>
<summary>
correction
</summary>

```{r, fig.width=18, fig.height=9}
ggplot(data=rnaseq2, mapping=aes(x=sample, y=log2_counts, fill=gene_biotype)) + 
  geom_boxplot()
```

</details>
<br>

4. Select only protein_coding and lincRNA and split again the boxes per gene_biotype.

<details>
<summary>
correction
</summary>

```{r}
filter(rnaseq2, gene_biotype=="protein_coding" | gene_biotype=="lincRNA") %>% ggplot(mapping=aes(x=sample, y=log2_counts, fill=gene_biotype)) + 
  geom_boxplot()
```

</details>

<br>

5. Add a **geom_violin()** layer. Set **alpha=0.3** in geom_violin. What is the alpha parameter?

<details>
<summary>
correction
</summary>

```{r}
filter(rnaseq2, gene_biotype=="protein_coding" | gene_biotype=="lincRNA") %>% ggplot(mapping=aes(x=sample, y=log2_counts, fill=gene_biotype)) + 
  geom_boxplot() +
  geom_violin(alpha=0.3)

# if boxplot and violin plots are misaligned, you can play with the position parameter in geom_violin, such as:
# geom_violin(position=position_dodge(0.7))
```

</details>

<br>

6. Look at the help page of **geom_boxplot()** and change the following parameters:

* Set outlier color to red
* Set outlier shape as triangles

<details>
<summary>
correction
</summary>

```{r}
filter(rnaseq2, gene_biotype=="protein_coding" | gene_biotype=="lincRNA") %>% ggplot(mapping=aes(x=sample, y=log2_counts, fill=gene_biotype)) + 
  geom_boxplot(outlier.colour = "red", outlier.shape = "triangle")
```

</details>

